---
title: "R Notebook"
output: html_notebook
---

### Degree distribution

```{r}
rm(list=ls())
library(networkdata)
library(intergraph)
library(network)
library(igraph)

# Load data
data(protein)
protein = simplify(protein, remove.multiple = TRUE, remove.loops = FALSE)
```

```{r}
# Select vertices with different lethality name
others = which(!(igraph::vertex_attr(protein, "lethality") %in% c("Unknown", "Lethal", "Non-Lethal")))

# Set selected vertice attributes to "Other"
protein = set_vertex_attr(protein, "lethality", others, "Other")

# Save a network object of protein dataset as well
protein_net = asNetwork(protein)
```

Let's visualize the degree distribution of the protein dataset.
```{r}
library(ggplot2)
protein.degrees.df = data.frame(degree = degree(protein))

ggplot(protein.degrees.df) + 
  ggtitle("Degree distribution of protein dataset") + 
  aes(degree) + geom_histogram(binwidth = 2)
```

The protein structure of yeast contains a main graph where most proteins interact and lots of proteins that have no interaction whatsoever. Let's take a look at the main graph
```{r}
dprotein <- decompose.graph(protein, mode="strong", min.vertices = 2)
main.interaction = dprotein[[1]]

l = layout_with_fr(main.interaction)
deg = degree(main.interaction, mode="all")
plot(main.interaction, layout=layout_with_lgl, vertex.size=deg/5, vertex.label=NA)
```

```{r}
main.interaction.degrees.df = data.frame(degree = degree(main.interaction))

ggplot(main.interaction.degrees.df) + 
  ggtitle("Degree distribution of main graph") + 
  aes(degree) + geom_histogram(binwidth = 2)
```
```{r}
summary(degree(main.interaction))
```

```{r}
N.v = as.numeric(vcount(main.interaction))
N.e = ecount(main.interaction)
```

```{r}
er.g = erdos.renyi.game(N.v, N.e, type="gnm")
ba.g =  barabasi.game(N.v)
```

```{r}
communities = fastgreedy.community(main.interaction)
communities
```

#### Small world
Does our main interaction graph has a small world property? In order to find that out we need to look at:
* Diameter (max length)
* Average shortest path length
* Clustering coefficient

```{r}
# Diameter
diameter(main.interaction)

# Average shortest path
sp = shortest.paths(main.interaction)
mean(sp[row(sp) != col(sp)])

# Clustering coefficient
transitivity(main.interaction)
```


These numbers don't tell us a lot, so we need to compare them to an for example random world.
```{r}
ntrials = 1000

num.dia.er = numeric(ntrials)
num.sp.er = numeric(ntrials)
num.clust.er = numeric(ntrials)

for (i in 1:ntrials) {
  er.g = erdos.renyi.game(N.v, N.e, type="gnm") 
  
  # Diameter
  num.dia.er[i] = diameter(er.g)
  
  # Shortest path (NA)
  sp = shortest.paths(er.g)
  sp[is.infinite(sp)] = NA
  num.sp.er[i] = mean(sp[row(sp) != col(sp)], na.rm = TRUE)
  
  # Clustering coefficient
  num.clust.er[i] = transitivity(er.g)
}
```

Diameter is longer than random graph. So is the clustering coefficient. We find that the mean shortest path of 

```{r}
sp = shortest.paths(main.interaction)

hist(num.sp.er)
abline(v = mean(sp[row(sp) != col(sp)]))

hist(num.clust.er, xlim = c(0,0.06))
abline(v=transitivity(main.interaction))

hist(num.dia.er, 10)
abline(v=diameter(main.interaction))
```
We have a lot more clusters than a random graph.

```{r}
ntrials = 1000
num.clust.er = numeric(ntrials)

for (i in 1:ntrials) {
  er.g = erdos.renyi.game(vcount(protein), ecount(protein), type="gnm") 
  
  num.clust.er[i] = clusters(er.g)$no
}

mean(num.clust.er)
clusters(protein)$no
```
WE HAVE A SMALL WORLD! WOO

