---
title: "Guillaume's part"
author: "Guillaume Franchi"
date: "4/1/2021"
output:
  html_document: default
  pdf_document: default
bibliography: bibliography.bibtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preprocessing

```{r, include = FALSE, echo = FALSE}
rm(list=ls())
library(network)
library(igraph)
library(intergraph)
library(networkdata)
library(sna)
library(ggplot2)
library(RColorBrewer)
```

```{r, echo = FALSE}
#-- load data and compute summary statistics
data(protein)
protein = simplify(protein, remove.multiple = TRUE)

# Select vertices with different lethality name
others = which(!(igraph::vertex_attr(protein, "lethality") %in% c("Unknown", "Lethal", "Non-Lethal")))

# Set selected vertice attributes to "Other"
protein = set_vertex_attr(protein, "lethality", others, "Other")

# Save a network object of protein dataset as well
protein_net = asNetwork(protein)
```

We take the giant component :

```{r,echo=FALSE}
prot_comp = component.largest(protein_net)
prot_small = protein_net[prot_comp,prot_comp]
smallnet = as.network(prot_small, directed=FALSE, remove.loops = FALSE )

lethality = network::get.vertex.attribute( protein_net, 'lethality')
sub_lethality = as.vector(subset(lethality, prot_comp!="FALSE"))

smallprot<-asIgraph(smallnet)
```

# Centrality

## Closeness centrality

```{r,echo=FALSE}
clo.cent<-igraph::closeness(smallprot,normalized = TRUE)
avg.clo.cent<-tapply(clo.cent, sub_lethality, mean)
avg.clo.cent
```
```{r,echo=FALSE}
hist(clo.cent)
```


## Betweenness centrality

```{r,echo=FALSE}
bet.cent<- igraph::betweenness(smallprot,normalized = TRUE)
avg.bet.cent<-tapply(bet.cent, sub_lethality, mean)
avg.bet.cent
```

```{r,echo=FALSE}
hist(bet.cent)
```

## Eigenvalue centrality

```{r,echo=FALSE}
eig.cent<- igraph::eigen_centrality(smallprot)$vector
avg.eig.cent<-tapply(eig.cent,sub_lethality,mean)
avg.eig.cent
```
```{r,echo=FALSE}
hist(eig.cent)
```


We cannot say that "lethal" proteins are more central than the others, and thus regarding each one of the centrality measure above.

# Clustering coefficient

We take a look at clustering coefficients in the network :

```{r}
cl<-transitivity(protein,type = "local")
```

```{r}
summary(cl)
```

We take the average of the clustering coefficient according to the different values of the 'lethality' attribute. First we remove the missing values :

```{r}
sub.cl<-cl[-which(is.na(cl))]
lethality.cl<-lethality[-which(is.na(cl))]
```



```{r}
avg.cl<-tapply(sub.cl,lethality.cl,mean)
avg.cl
```
Nodes with lethal attribute tends to have a slightly higher clustering coefficient. However it may be not relevant as the number of nodes with undefined clustering coefficient is very high...

# ERGM with `gwdegree()`

```{r,echo=FALSE}
library(ergm)
```
We want now to take into account the degree distribution and the effect of lethality value in our model.

First, we tune our parameter of decay for the `gwdegree()` function, between 0.1 and 0.9. We don't look for higher parameter, since the decrease of the degree distribution is fast, it seems reasonable to look for a small value of the decay parameter.

```{r,echo=FALSE,message=FALSE}
decay<-seq(0.1,0.9,by=0.1)
bic<-rep(NA,length(decay))

for (i in 1:length(decay)){
  fit<-ergm(protein_net ~ edges + nodefactor("lethality",base=2)+ gwdegree(decay[i],fixed=TRUE))
  bic[i]<-summary(fit)$bic
}
```


```{r}
min.decay<-decay[which.min(bic)]
ergm.fit.4 = ergm(protein_net ~ edges + nodefactor("lethality",base=2)  + gwdegree(min.decay,fixed=TRUE))
```
```{r}
summary(ergm.fit.4)
```

The BIC criterion is slightly better, so it seems a better fit than the other models. Let's check the goodness of fit with the following plot :

```{r}
gof.4<-gof(ergm.fit.4)
par(mfrow=c(2,2))
plot(gof.4)
```

Here, the minimum geodesic distance indicates a good fit of the model, and the transitivity is still well captured. However, we still don't modelze properly the degree distribution.

# ERGM with `gwesp()`

We finally try a model taking into account both the degree and the transitivity.

```{r}
ergm.fit.5<- ergm(protein_net ~ edges + nodefactor("lethality",base=2) + gwdegree(log(2),fixed=TRUE) +
  gwesp(log(3),fixed=TRUE))
```
```{r}
summary(ergm.fit.5)
```
```{r}
gof.5<-gof(ergm.fit.5)
par(mfrow=c(2,2))
plot(gof.5)
```



